---
title: "Covid91 vaccine study Final Spring 2023"
author: "Roxan Barakat"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalFall2022.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```


# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of sub populations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.



# Coding details 

First, the variables from FinalData needed to be changed from character variables into factor variables. The appropriate dplyr code was used to do this. The following code was used to achieve this: 

FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

Next, specific subgroups had to be separated from the broader categories in order to properly studied. Male and Female were separated into subgroups, as well as sexual orientation and drug use. All of these categories are isolated to see whether or not these particular populations should or should not take the  new covid291 vaccine.


In order to create the effectiveness interval for all four categories, each group of people in this study had to be stored in separate variables. 


# Males

## Barplot
```{r}
dd2 <- male %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of males and the relationship to the contraction of covid921 by either whether they were given the actual vaccine or the placebo in its place. 

There are two plots from this chunk of coding. In first plot,the x-axis represents infected,in terms of whether the not participants were infected with covid921. The y-axis represents the count, or number of people that got covid921 or not. In the second plot, the one above, the left two bars of the graph are all the people that got covid921. The x-axis represents whether a person got covid921 or not. The y-axis, however, instead of representing the number of people who got covid921, now represents the percentage of people that got covid921 after taking the the placebo or taking the Vaccine. 

The plot above shows us that there was about 68% of the male population who were given the placebo and had in fact contracted covid921, while around 49% of males who were also given the placebo were okay. From the data above, we can see that there was about 32% of males who were given the vaccine and still contracted covid921, while around 51% of the male population was given the vaccine and did not contract covid921. Overall, there was a substantially higher percentage of people who took the placebo contracted covid921 versus those who were given the vaccine. The vaccine appears to be effective in protection against covid921. The odds of getting covid921 while taking the vaccine are 1:2. The men that took the placebo were two times more likely to get covid921 than those taking the vaccine.

The codes that create this plot start with the code block beginning with dd2. The code dd2 can be seen in the environment pane and if inspected contains 4 variables. These variables include infected, treatment, count, and prcnt (which stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot. For both basicC and basicCC these variables include data, layers, scales, mapping, theme, coordinates, facet, plot_env, labels. 


## Numerical results
```{r}
table1 <- xtabs(~infected + treatment, data=male)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of males that got covid921 after being given the placebo and the vaccine. Among the males, 32% got covid921 after taking the vaccine, while the approximately 68% of males that were given the placebo contracted covid921. The vaccine is clearly effective in preventing covid921 in males and it is recommended. Those who took the placebo were much more likely to contract covid921. The odds of getting covid921 while taking the vaccine are 1:2. The men that were given the placebo were two times more likely to get covid921 than those who were given the vaccine.

##  Inferential 
```{r}
chisq.test(table1)
chisqtestGC(table1)
malefish <- fisher.test(table1)
malefish
```
```{r}
effect <- malefish$p.value/(malefish$p.value +1)
effectivenessmales = malefish$estimate/(malefish$estimate +1)*100
print(effectivenessmales)
```
```{r}
low=(malefish$conf.int[1]/(malefish$conf.int[1] +1))*100
hi=(malefish$conf.int[2]/(malefish$conf.int[2] +1))*100
options(digits=2)
```

From the data above the information we are focusing on the most are the results we gathered from the Fisher's Exact test. The p-value displayed above is extremely small with the value of 0.000000000002. This is an awfully small number and seeing this number means that it is very unlikely that these results will occur by chance.

The null hypothesis predicted that the deaths related to covid921 are independent of sex. Since it is clear that sex does play a role in the death rate of individuals, the alternative hypothesis, that there is a correlation between covid921 deaths and sex, is the favorable option.

The odd ratio above illustrates a ratio of 2.2. This means that the males who received the placebo were two times more likely to get covid921 than those who were given the vaccine. 


## Effectiveness Confidence Interval  `r low`  to `r hi`

Above is the effectiveness for the covid921 vaccine for the male population. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is effective in protecting males from contracting covid921. 

The odds ratio is approximately how effective the vaccine is. The odds ratio for the male population can be seen above as 69. The effectiveness confidence interval for the male population is between 63.40 to 73.35.


# Females

## Barplot
```{r}
dd2 <- female %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of females and the relationship to the contraction of covid921 by either whether they were given the actual vaccine or the placebo in its place. 

There are two plots from this chunk of coding. In first plot,the x-axis represents infected,in terms of whether the not participants were infected with covid921. The y-axis represents the count, or number of people that got covid921 or not. In the second plot, the one above, the left two bars of the graph are all the people that got covid921. The x-axis represents whether a person got covid921 or not. The y-axis, however, instead of representing the number of people who got covid921, now represents the percentage of people that got covid921 after taking the the placebo or taking the Vaccine. 

In this plot it contrasts the previous and shows that there was about 42% of the female population who were given the placebo and had in fact contracted covid921, while around 50% of females who were also given the placebo were okay. From the data above, we can see that there was about 57% of females who were given the vaccine and still contracted covid921, while around 49% of the female population was given the vaccine and did not contract covid921. Overall, there was a substantially higher percentage of people who took the vaccine and contracted covid921 versus those who were given the placebo. The vaccine appears to be ineffective in protection against covid921. The odds of getting covid921 while taking the vaccine are 3:2. The females that took the vaccine were one and a half times more likely to get covid921 than those taking the placebo.

The codes that create this plot start with the code block beginning with dd2. The code dd2 can be seen in the environment pane and if inspected contains 4 variables. These variables include infected, treatment, count, and prcnt (which stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot. For both basicC and basicCC these variables include data, layers, scales, mapping, theme, coordinates, facet, plot_env, labels.

## Numerical results
```{r}
table1 <- xtabs(~infected + treatment, data=female)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of females that got covid921 after being given the placebo and the vaccine. Among the females, 57% got covid921 after taking the vaccine, while the approximately 43% of females that were given the placebo contracted covid921. The vaccine is clearly ineffective in preventing covid921 in females. Those who took the vaccine were much more likely to contract covid921. The odds of getting covid921 while taking the vaccine are 3:2. The females that were given the vaccine were one and half times more likely to get covid921 than those who were given the placebo.


##  Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
femalefish <- fisher.test(table1)
femalefish
```


```{r}
effect <- femalefish$p.value/(femalefish$p.value +1)
effectivenessfemale = femalefish$estimate/(femalefish$estimate +1)*100
print(effectivenessfemale)
```

```{r}
low=(femalefish$conf.int[1]/(femalefish$conf.int[1] +1))*100
hi=(femalefish$conf.int[2]/(femalefish$conf.int[2] +1))*100
options(digits=2)
```

From the data above the information we are focusing on the most are the results we gathered from the Fisher's Exact test. The p-value displayed above is small with the value of 0.005. This is small number  means that it is very unlikely that these results will occur by chance.

The null hypothesis predicted that the deaths related to covid921 are independent of sex. Since it is clear that sex does play a role in the death rate of individuals, the alternative hypothesis, that there is a correlation between covid921 deaths and sex, is the favorable option.

The odd ratio above illustrates a ratio of 0.74. This means that the females who received the placebo were one and half times more likely to get covid921 than those who were given the vaccine. 



## Effectiveness Confidence Interval 


# LGBTQ

## Barplot

## Numerical results

##  Inferential 

## Effectiveness Confidence Interval 


# Druggies

## Barplot

## Numerical results

##  Inferential 

## Effectiveness Confidence Interval


# Overall Results and Conclusions